---
title: "Store Matching Using Fuzzyjoin In R"
description: |
  Matching stores between two datasets.
author:
  - name: Zahier Nasrudin
    url: https://zahier-nasrudin.netlify.app/
date: 2020-09-10
preview: images/market.jpg
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = F, message = F)
```

# Introduction

Last week, we had a request from a client. We were given a list of 600 stores and I needed to match those stores with the stores that we have in our database.

So, I used "fuzzyjoin" package in R to find the similarity between these stores based on the address list given.

For this project, I am not going to use the dataset given by the client (confidential). Instead, I am going to use a dataset available in [Kaggle.](https://www.kaggle.com/ben1989/target-store-dataset)

# Dataset

It is a list of Target stores (as of April 2017). It has a lot of useful attributes such as operating hour, open date, and latitude & longitude. However, I will only be requiring the address column for this project:

```{r}

library(tidyverse)
library(janitor)
library(readr)
library(fuzzyjoin)
library(here)
 
targets <- read_csv(here("_posts/fuzzy-matching/1. Input Files/targets.csv")) %>%
  clean_names() %>%
  select(id, address_formatted_address)
```

I will copy this data set, but this time I will remove the vowels, and the id column

```{r}
targets_missing <- targets %>%
  select(-id) %>%
  mutate(address_formatted_address = str_remove_all(
    tolower(address_formatted_address), 
    "[aeiou]"))

```

Next, I will use stringdist\_left\_join() to find the stores between the targets data frame and "targets\_missing" data frame. I will then use slice\_min to get **the minimum value (distance score) for each of the store** because **I just want only a match per store**, and **I want it to be the closest distance**. (Hence, the minimum value per store).

```{r}
stores_matching <- targets %>%
  stringdist_left_join(targets_missing, by = "address_formatted_address",
                       method = "cosine", distance_col = "score")
 
stores_matching <- stores_matching %>%
  group_by(address_formatted_address.x) %>%
  slice_min(order_by = score)

```

```{r}
stores_matching %>%
  head() %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling()
```

# Closing

Now, this method definitely helped me when I had to find at least 600 stores. And no, I definitely did not consider to do it manually.

Please let me know what you think of this and what you could have done differently. Thanks for reading and have a nice day

