---
title: "Covid 19: Malaysia"
description: |
  Analyzing the number of Covid 19 cases, number of tests, and clusters in Malaysia
author:
  - name: Zahier Nasrudin
    url: https://zahier-nasrudin.netlify.app/
date: 08-09-2021
preview: images/scientist.jpg
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(skimr)
library(janitor)
library(here)
library(reactable)
library(reactablefmtr)

theme_set(theme_minimal())

```


```{r data-loading}

no_cases <- read_csv(("https://raw.githubusercontent.com/MoH-Malaysia/covid19-public/main/epidemic/cases_state.csv")) %>%
  clean_names()

no_tests <- read_csv("https://raw.githubusercontent.com/MoH-Malaysia/covid19-public/main/epidemic/tests_malaysia.csv") %>%
  clean_names()
 
no_deaths <- read_csv("https://raw.githubusercontent.com/MoH-Malaysia/covid19-public/main/epidemic/deaths_state.csv") %>%
  clean_names()

cluster <- read_csv("https://raw.githubusercontent.com/MoH-Malaysia/covid19-public/main/epidemic/clusters.csv") %>%
  clean_names() %>%
  mutate(category =  str_to_title(category))



```


## Description

Malaysia's Ministry of Health (MoH) just released a public MoH Github repository. The github can be found here: https://github.com/MoH-Malaysia/covid19-public. So it would be nice if I could explore some of the datasets

## Analysis

```{r}

cases_deaths <- no_cases %>%
  inner_join(no_deaths, by = c("date", "state")) %>%
  mutate(cases_new = abs(cases_new))

no_cases_malaysia <- no_cases %>%
  group_by(date) %>%
  summarise(cases_new = sum(cases_new)) %>%
  ungroup()

no_tests_revised <- no_tests %>%
  filter(date != "2020-03-16") %>%
  inner_join(no_cases_malaysia, by = "date") %>%
  mutate(total_test = rtk_ag + pcr) %>%
  mutate(ratio_covid = cases_new / total_test)



```

### Number of cases

Here is an overview of the number of cases in Malaysia by (State).

```{r, fig.width= 10, fig.height= 7}

cases_deaths %>%
  ggplot(aes(x = date, y = cases_new)) + 
  geom_line(colour = "light blue") +
  scale_x_date(date_labels =  "%b %Y",
                         date_breaks = "3 months") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma) +
  facet_wrap(~ state, scales = "free_y") + 
  labs(title = "Number of Covid 19 cases",
       subtitle = "17th of March 2020 - 8th of August 2021",
       x = "",
       y = "")


#ggsave("covidcases.png", width = 9)

```

### Number of deaths

The number of deaths in Malaysia by (State).

```{r, fig.width= 10, fig.height= 7}

cases_deaths %>%
  group_by(state) %>%
  summarise(deaths_new = sum(deaths_new)) %>%
  mutate(state = fct_reorder(state, deaths_new)) %>%
  ggplot(aes(x = state, y = deaths_new)) + 
  geom_col(fill = "light blue") +
  geom_text(aes(label=  scales::comma(deaths_new, accuracy = 1)), 
                vjust=0.5, hjust=0, size = 3.2)+
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
    labs(title = "Number of Covid-19 deaths",
       subtitle = "17th of March 2020 - 8th of August 2021",
       x = "",
       y = "")


#ggsave("coviddeaths.png", width = 9)


```


### Ratio (Positive Covid-19 tests)

The data released for this is not by state level. Hence, we can only compare the ratio (number of Covid 19 cases / number of tests) by total country. The total tests variable is a combination of  Antigen Rapid Test Kits (RTK-Ag) and Real-time Reverse Transcription Polymerase Chain Reaction (RT-PCR) technology

```{r, fig.width= 10, fig.height= 7}


no_tests_revised %>%
  ggplot(aes(x = date, y = ratio_covid)) +
  geom_line(color = "light blue", size = 0.7) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_date(date_labels =  "%b %Y",
                         date_breaks = "3 months") +
  labs(title = "Percentage of positive Covid 19 tests",
       subtitle = "17th of March 2020 - 8th of August 2021",
       x = "",
       y = "No. of Covid 19 cases / No. of tests")


#ggsave("percentagepositive.png", width = 9)

```


### Cluster

Lastly, exploring the cluster groups in Malaysia. By focusing on only cluster groups that are still active, workplace is the most common Covid 19-cluster category:


```{r}


cluster_table <- cluster %>%
  filter(status == "active") %>%
  group_by(category) %>%
  summarise(cases_total = sum(cases_total),
            icu = sum(icu),
            deaths = sum(deaths),
            recovered = sum(recovered),
            clustergroup_total = length(cluster)) %>%
  relocate(clustergroup_total, .before = cases_total) %>%
  arrange(desc(cases_total)) %>%
  reactable(
    columns = list(
    category = colDef(name = "Category"),
    cases_total = colDef(name = "Total Cases",
                                cell = data_bars(., 
                                               fill_color = c("light blue"),
                                               number_fmt = scales::comma)),
    clustergroup_total = colDef(name = "Number of cluster groups",
                                cell = data_bars(., 
                                               fill_color = c("light blue"),
                                               number_fmt = scales::comma)),
    icu = colDef(name = "ICU",
                 cell = data_bars(.,
                                  fill_color = c("light blue"),
                                  number_fmt = scales::comma)),
    deaths = colDef(name = "Deaths", 
                    cell = data_bars(., 
                                     fill_color = c("light blue"),
                                     number_fmt = scales::comma)),
    recovered = colDef(name = "Recovered",
                       cell = data_bars(., 
                                     fill_color = c("light blue"),
                                     number_fmt = scales::comma))
  )) %>%
  add_title("Covid-19 Clusters (Active cases)", font_family = "Proza Libre", font_size = 25) %>%
  add_subtitle("1st of March 2020 - 8th of August 2021",  font_family = "Proza Libre", font_size = 15)


cluster_table

#save_reactable(cluster_table, "table.png")

```

Thanks again to MoH for releasing Covid-19 related datasets!

