---
title: "malaytextr: Stemming malay text"
description: |
  Stemming malay text using stem_malay() from malaytextr package in R  
author:
  - name: Zahier Nasrudin
    url: https://zahier-nasrudin.netlify.app/
date: "`r Sys.Date()`"
preview: images/malaysiaflag.jpg
bibliography: citations.bib
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)



```

# Introduction

A package for handling malay text in R is currently being developed ([malaytextr](https://github.com/zahiernasrudin/malaytextr)). It is aimed to easily handle malay text (stemming, stopwords and more).

As of now:

1.  malayrootwords: A malay root words data frame: Can be served as a dictionary/reference
2.  stem_malay(): A function to stem malay words. For example, "semangatnya" will now change to "semangat". This function will first find the malay words in a dictionary (malayrootwords can be used). It will then follow the framework from [@khanmalay]

# Example

### Install malaytextr package

The malaytextr package can be installed from Github:

```{r, eval=FALSE}

# install.packages("devtools")
devtools::install_github("zahiernasrudin/malaytextr")


```

### Load libraries

```{r}

library(malaytextr)
library(dplyr) # for data wrangling

```

### Malay root words

There is a data frame (malayrootwords) that can be used as a dictionary. This data frame will definitely be updated with more words in the near future. Currently, the malayrootwords data frame contains 4244 rows:

```{r}

malayrootwords

```

### Stemming

To perform stem_malay(), it is pretty simple to do it. For example you want stem "banyaknya". You will just need to specify "banyaknya" and our dictionary which is "malayrootwords". It will then return a data frame ("banyaknya" and the root word "banyak"):

```{r}

stem_malay(Word = "banyaknya",
           dictionary = malayrootwords)


```

If you want to stem more than one word, for example "banyaknya" , "sakitnya" and "kemahuan":

```{r}

stem_malay(Word = c("banyaknya", "sakitnya", "kemahuan"),
           dictionary = malayrootwords)


```

Now, if you have a data frame that contains word that you want to stem:

1.  Create a data frame called x to be used as an example
2.  Specify the column you want to stem in "col_feature1"

```{r}

x <- data.frame(text = c("banyaknya","sangat","terkedu", "pengetahuan"))

stem_malay(Word = x, 
           dictionary = malayrootwords, 
           col_feature1 = "text")

```

Please have a look and to report a bug, please file an issue on [Github](https://github.com/zahiernasrudin/malaytextr/issues)

Thank you :)
