---
title: " Working with date-time data: Lubridate & Janitor"
description: |
  Handling date-time data using Lubridate package & Janitor package in R-Studio
author:
  - name: Zahier Nasrudin
    url: https://zahier-nasrudin.netlify.app/
date: 06-02-2021
preview: images/calendar.jpg
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)

```

# Introduction

Dealing with date-time is something that I will usually avoid when using R. One of the reasons:

1.  It can be frustrating: Due to the fact that there could be multiple types of date-time, **or** different languages (locale)

However, I had to deal with a few date-time datasets recently. By googling & researching, I found lubridate & janitor packages to be quite interesting to use. Two resources that I have been using & like:

1.  [R for Data Science](https://r4ds.had.co.nz/dates-and-times.html)
2.  [Lubridate package](https://lubridate.tidyverse.org/)
3.  [Janitor package](https://rdrr.io/cran/janitor/man/excel_numeric_to_date.html)

## Data Loading

```{r data-loading}

# Libraries required

library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(lubridate)


date_time <- read_excel(here("_posts/work-date-time/date-time.xlsx"))



```

## Description

The dataset used in this blog contains 2 columns & 16,254 records :

```{r}

glimpse(date_time)



```

1.  Date: Date of transaction made
2.  Timestamp: Time of transaction made

## Exploring Data

Looking at the structure of the dataset above, there are 2 problems.

1.  Date column is not displaying properly
2.  Timestamp column is not displaying properly. For example, 17:39:29 is displayed as 1899-12-31 17:39:29 when uploaded in R. Therefore we will need to extract just the timestamp later on

### First problem: Date column

Lets have a look at the first problem:

1.  Weird numeric format (44287, 44310, 44311, 44312). This happens because the date is in excel date format (it counts days). A similar problem can found in [stack overflow](https://stackoverflow.com/questions/43230470/how-to-convert-excel-date-format-to-proper-date-in-r/62334132)
2.  Multiple date formats in a column (04.01.2021, 1-Mei-2021, 27/5/2021)

```{r}

date_time %>%
  count(Date, sort = TRUE)

```

Fixing the numeric format to date is easy to solve using excel_numeric_to_date using the janitor package:

```{r}

date_time_processed <- date_time %>%
  mutate(`Date 2` = excel_numeric_to_date(as.numeric(Date)))

```

So we managed to convert from numeric to date format (Compare Date column vs Date 2 column):

```{r}

date_time_processed %>%
  count(Date, `Date 2`, sort = TRUE )


```

Now we need to standardize the different date formats (04.01.2021, 1-Mei-2021 and 27/5/2021). To fix that, I will be using parse_date_time2 from lubridate package. Once the date is extracted, I will then convert it to date format. The process:

1.  Using if_else() to find the missing values in "Date 2" column. For this case, the ones that failed to convert from numeric to date (04.01.2021, 1-Mei-2021 and 27/5/2021)
2.  Parse 04.01.2021, 1-Mei-2021 and 27/5/2021 using parse_date_time2 and specify the pattern (%d%m%y) which is day, month and year.
3.  Convert them to date format using as_date

```{r}

date_time_processed <- date_time_processed %>%
  mutate(`Date 3` = if_else(is.na(`Date 2`),
                            parse_date_time2(Date, "%d%m%Y") %>% as_date(),
                            `Date 2`)) 

```

Checking it back, we managed to convert all them successfully (Date 3):

```{r}

date_time_processed %>%
  count(Date, `Date 2` , `Date 3`)
```

### Second problem: Timestamp column

Now for the last problem, timestamp! This is pretty straightforward. Here is the process:

1.  Extract the timestamp using format() using pattern "%H:%M:%S" which is hour, minute and second
2.  I will then combine the Date 3 column with extracted Timestamp2 column to create a date-time column

```{r}

date_time_processed <- date_time_processed %>%
  mutate(Timestamp2 = format(Timestamp, "%H:%M:%S"),
         Final_date = as_datetime(paste(`Date 3`, Timestamp2)))
```

The first 6 rows of our final_date column:

```{r}

date_time_processed %>%
  select(Final_date) %>%
  head()

```

Thanks for reading!
